---
title: "Bioconductor in R Universe"
format:
  html:
    code-fold: true
editor: visual
engine: knitr
---

Go to [windows](windows.qmd).

Using the R Universe API, this document looks at differences in packages between
R Universe and Bioconductor to identify potential issues that require a fix and
a processes. In particular, we compare

- Bioconductor 3.23 and https://bioc.r-universe.dev
- Bioconductor 3.22 and https://bioc-release.r-universe.dev

In https://bioc.r-universe.dev, there are versions that differ by 2 `y` versions.
How did this happen, how can we fix it, and make sure it always follows the
versions that should be associated with the devel branch of Bioconductor?

There are packages not in R Universe due to early errors. Who is responsible for
addressing these (maintainer or core) and what's the process?

There are also some packages named as remotes that seem accessible via the API,
which probably just needs a fix.

```{r install_universe, echo = FALSE, message = FALSE}
library(BiocPkgTools)
library(dplyr)
library(universe)
library(biocUniTools)
library(httr2)
library(reactable)

.LIMIT <- 3000L
options(max.print = .LIMIT)
```

## R Universe and Bioconductor 3.23

### Bioc 3.23: Not in Bioconductor

These packages seem available via the API, e.g.,
https://bioc.r-universe.dev/api/packages/ggpattern, but it doesn't appear on
https://bioc.r-universe.dev/apis. Their repositories have been set as remotes
https://github.com/r-universe/bioc/blob/master/.remotes.json, like for ALDEx_bioc: https://github.com/ggloor/ALDEx_bioc/blob/master/DESCRIPTION. Maybe they should
not be accessible via the API.

```{r not-in-bbs-devel, echo = FALSE}
r_version <- "4.6.0"
bioc_version <- "3.23"
uni <- "bioc"
uni_os_branch <- "devel"
ru_builds <- biocUniTools::get_uni_pkgs(uni, uni_os_branch, r_version,
                                        "linux", verbose = FALSE)
bbs_pkgs_i <- BiocPkgTools::biocBuildReport(version = bioc_version,
                                            pkgType = "software")
bbs_pkgs <- unlist(unique(bbs_pkgs_i$pkg))

not_in_bbs <- setdiff(ru_builds$Package, bbs_pkgs) 
if (length(not_in_bbs) > 0) {
    not_in_bbs
}
```


```{r not-in-ru-details-devel, echo = FALSE}
bbs_builds <- bbs_pkgs_i |>
    dplyr::filter(node == "nebbiolo1", stage == "checksrc") |>
    dplyr::select(pkg, version, Deprecated, git_last_commit, result) |>
    dplyr::rename(Package = pkg, Version = version, BBS_commit = git_last_commit,
                  Check = result)
ru_builds <- biocUniTools::get_uni_pkgs(uni, uni_os_branch, r_version, "linux")

builds <- merge(bbs_builds, ru_builds, by = c("Package", "Version"))
unmatched_ru_builds <- ru_builds |>
    dplyr::filter(!Package %in% builds$Package) |>
    dplyr::select(Package, Version, RU_commit) |>
    dplyr::rename(RU_Version = Version)
unmatched_bbs_builds <- bbs_builds |>
    dplyr::filter(!Package %in% builds$Package) |>
    dplyr::select(Package, Version, BBS_commit, Deprecated) |>
    dplyr::rename(BBS_Version = Version)
unmatched <- dplyr::full_join(unmatched_bbs_builds, unmatched_ru_builds,
                              by = "Package")
not_in_bbs_pkgs <- setdiff(unmatched$Package, unmatched_bbs_builds$Package)
not_in_ru_pkgs <- data.frame(Package = setdiff(unmatched$Package, unmatched_ru_builds$Package),
                             Buildurl = "")
for (pkg in not_in_ru_pkgs$Package) {
    u <-paste0("https://", uni, ".r-universe.dev/api/packages/", pkg)
    pkglist <- httr2::request(u) |>
        httr2::req_perform() |>
        httr2::resp_body_json()
    not_in_ru_pkgs[not_in_ru_pkgs$Package == pkg, ]$Buildurl <- pkglist$`_buildurl`
}
```

### Bioc 3.23: Early failures and missing dependencies

These packages fail early, sometimes due to missing dependencies or possibly
DESCRIPTION file errors.

```{r missing-deps-or-early-failures-devel, echo = FALSE}
reactable::reactable(not_in_ru_pkgs, defaultPageSize = 50, compact = TRUE,
                     columns = list(
                        Buildurl = reactable::colDef(cell = function(value, index) {
                            htmltools::tags$a(href = not_in_ru_pkgs[index, "Buildurl"],
                                              target = "_blank",
                                              as.character(value))
                            }))
)
```

### Bioc 3.23: Different versions

We should update the packages where y differs. With https://bioc.r-universe.dev,
y sometimes differs by 2. How does this happen and how do we fix it?

```{r different-versions-devel, echo = FALSE}
unmatched <- unmatched |>
    dplyr::filter(!Package %in% c(not_in_ru_pkgs$Package, not_in_bbs_pkgs)) |>
    dplyr::select(Package, BBS_Version, RU_Version, BBS_commit, RU_commit,
                  Deprecated)
reactable::reactable(unmatched, searchable = TRUE, showPageSizeOptions = TRUE,
                     compact = TRUE)
```
## Bioc 3.23: Same versions

These are probably ok. Only showing result of linux builds. `Check` is the BBS
result.

```{r same-versions-devel, echo = FALSE}

condensed_builds <- builds |>
    dplyr::mutate(CommitMatch = BBS_commit == RU_commit) |>
    dplyr::select(Package, Version, Deprecated, CommitMatch, Check, JobCheck,
                  BinariesCheck, JobUrl, BinariesUrl)
reactable::reactable(condensed_builds, searchable = TRUE,
                     showPageSizeOptions = TRUE, compact = TRUE,
                     columns = list(
                        JobCheck = reactable::colDef(cell = function(value, index) {
                            htmltools::tags$a(href = condensed_builds[index, "JobUrl"],
                                              target = "_blank",
                                              as.character(value))
                            }),
                        BinariesCheck = reactable::colDef(cell = function(value, index) {
                            htmltools::tags$a(href = condensed_builds[index, "BinariesUrl"],
                                              target = "_blank",
                                              as.character(value))
                            }),
                        JobUrl = reactable::colDef(show = FALSE),
                        BinariesUrl = reactable::colDef(show = FALSE)
                         )
)
```

## R Universe and Bioconductor 3.22

### Bioc 3.22: Not in BBS

```{r not-in-bbs-release, echo = FALSE}
r_version <- "4.5.0"
bioc_version <- "3.22"
uni_os_branch <- "release"
uni <- "bioc-release"
ru_builds <- biocUniTools::get_uni_pkgs(uni, uni_os_branch, r_version,
                                        "linux", verbose = FALSE)
bbs_pkgs_i <- BiocPkgTools::biocBuildReport(version = bioc_version,
                                            pkgType = "software")
bbs_pkgs <- unlist(unique(bbs_pkgs_i$pkg))
not_in_bbs <- setdiff(ru_builds$Package, bbs_pkgs) 
if (length(not_in_bbs) > 0) {
    not_in_bbs
}
```

```{r not-in-ru-details-release, echo = FALSE}
bbs_builds <- bbs_pkgs_i |>
    dplyr::filter(node == "nebbiolo2", stage == "checksrc") |>
    dplyr::select(pkg, version, Deprecated, git_last_commit, result) |>
    dplyr::rename(Package = pkg, Version = version, BBS_commit = git_last_commit,
                  Check = result)

builds <- merge(bbs_builds, ru_builds, by = c("Package", "Version"))
unmatched_ru_builds <- ru_builds |>
    dplyr::filter(!Package %in% builds$Package) |>
    dplyr::select(Package, Version, RU_commit) |>
    dplyr::rename(RU_Version = Version)
unmatched_bbs_builds <- bbs_builds |>
    dplyr::filter(!Package %in% builds$Package) |>
    dplyr::select(Package, Version, BBS_commit, Deprecated) |>
    dplyr::rename(BBS_Version = Version)
unmatched <- dplyr::full_join(unmatched_bbs_builds, unmatched_ru_builds, by = "Package")
not_in_bbs_pkgs <- setdiff(unmatched$Package, unmatched_bbs_builds$Package)
not_in_ru_pkgs <- data.frame(Package = setdiff(unmatched$Package, unmatched_ru_builds$Package),
                             Buildurl = "")
for (pkg in not_in_ru_pkgs$Package) {
    u <-paste0("https://", uni, ".r-universe.dev/api/packages/", pkg)
    pkglist <- httr2::request(u) |>
        httr2::req_perform() |>
        httr2::resp_body_json()
    not_in_ru_pkgs[not_in_ru_pkgs$Package == pkg, ]$Buildurl <- pkglist$`_buildurl`
}
```

### Bioc 3.22: Early failures and missing dependencies

```{r missing-deps-or-early-failures-release, echo = FALSE}
reactable::reactable(not_in_ru_pkgs, defaultPageSize = 50, compact = TRUE,
                     columns = list(
                        Buildurl = reactable::colDef(cell = function(value, index) {
                            htmltools::tags$a(href = not_in_ru_pkgs[index, "Buildurl"],
                                              target = "_blank",
                                              as.character(value))
                            }))
)
```

### Bioc 3.22: Different versions

Better: version differences in only z.

```{r different-versions-release, echo = FALSE}
unmatched <- unmatched |>
    dplyr::filter(!Package %in% c(not_in_ru_pkgs$Package, not_in_bbs_pkgs)) |>
    dplyr::select(Package, BBS_Version, RU_Version, BBS_commit, RU_commit,
                  Deprecated)
reactable::reactable(unmatched, searchable = TRUE, showPageSizeOptions = TRUE,
                     compact = TRUE)
```

### Bioc 3.22: Same versions

Only showing result of linux builds and `Check` is the BBS result.

```{r same-versions-release, echo = FALSE}
condensed_builds <- builds |>
    dplyr::mutate(CommitMatch = BBS_commit == RU_commit) |>
    dplyr::select(Package, Version, Deprecated, CommitMatch, Check, JobCheck, BinariesCheck,
                  JobUrl, BinariesUrl)
reactable::reactable(condensed_builds, searchable = TRUE,
                     showPageSizeOptions = TRUE, compact = TRUE,
                     columns = list(
                        JobCheck = reactable::colDef(cell = function(value, index) {
                            htmltools::tags$a(href = condensed_builds[index, "JobUrl"],
                                              target = "_blank",
                                              as.character(value))
                            }),
                        BinariesCheck = reactable::colDef(cell = function(value, index) {
                            htmltools::tags$a(href = condensed_builds[index, "BinariesUrl"],
                                              target = "_blank",
                                              as.character(value))
                            }),
                        JobUrl = reactable::colDef(show = FALSE),
                        BinariesUrl = reactable::colDef(show = FALSE)
                         )
)
```

```{r sessionInfo}
sessionInfo()
```
